---
title: "AI Predict"
author: Tony Boyles (Anthony@Boyles.cc)
date: April 12, 2016
output:
  md_document:
    variant: markdown_github
bibliography: bibliography.bib
---

```{r echo=FALSE}
knitr::opts_chunk$set(
  comment=NA,
  fig.width=10,
  fig.height=6
)
```

An R Package of datasets to help predict the timeline preceding an Intelligence Explosion.

## Installation

```{r, eval=FALSE}
install.packages("https://github.com/AABoyles/AIPredict/archive/master.tar.gz", type = "source")
```

## Data

This package contains the following datasets:

 * `ai_moores_law` - contains observations contributing to estimates of Moore's law.
 * `ai_koomeys_law` - contains observations contributing to estimates of Koomey's law. Derived from the dataset constructed by Jonathan Koomey, which is aarchived in the `data-raw/` directory.
 * `ai_prediction` - contains public estimates of Artificial Intelligence milestones. Derived from the dataset produced by [AI Impacts](http://aiimpacts.org/miri-ai-predictions-dataset/), which is archived in the `data-raw/` directory.
 * `ai_bitcoin_hashrate` - contains the instantaneous hashrate of the bitcoin network measured daily at 6:15:05pm UTC.
 * `ai_animal_neurons` - contains the merged tables in the Wikipedia Article, [List of Animals by Number of Neurons](https://en.wikipedia.org/w/index.php?title=List_of_animals_by_number_of_neurons&oldid=710786267)
 * `ai_fli_winners` - contains the published data of the winners of the Future of Life Institute's 2015 RFP for grants on research for safe artificial intelligence.

### Moore's Law

Roughly stated, Moore's law predicts that the density of transistors in a single processor core grows exponentially. [@Moore1998-ze] It is a widely used and cited metric in predictions about the development of Artificial General Intelligence. Perhaps the best-known of these is futurist Ray Kurzweil's projections in The Age of Spiritual Machines and The Singularity is Near, which are based on simple of extrapolations of Moore's law.

```{r, message=FALSE}
library(AIPredict)
library(dplyr)
library(ggplot2)

ai_moores_law %>%
  ggplot(aes(Year, Transistors)) +
  geom_point() +
  scale_y_log10() +
  stat_smooth(method="lm")
```

### Koomey's Law

Koomey's law states that the electricty required to execute some number of computations declines exponentially over time [@Koomey2011-fc]. While less well-known than Moore's law, it offers another critical benchmark for comparison to the human brain.

```{r, message=FALSE}
ai_koomeys_law %>%
  ggplot(aes(Year, WattsPerMCPS)) +
  geom_point() +
  scale_y_log10() +
  stat_smooth(method="lm")
```

Every decade, the energy cost of computing falls approximately two orders of magnitude.

### Top 500 Supercomputers

The Top 500 Supercomputers. Another good source for demonstrating Moore's and Koomey's laws.

```{r}
ai_top500 %>%
  filter(Rank == 1) %>%
  mutate(RMAX=ifelse(is.na(RMax), Rmax, RMax)) %>%
  filter(!is.na(RMAX)) %>%
  ggplot(aes(Year, RMAX)) +
  geom_jitter(width = 1, height = 0, alpha=.3, size=5) +
  scale_y_log10() +
  stat_smooth(method = "lm")
```

While generally accessible, these data currently require a good deal of cleaning, which I'll perform and document in due course.

### Bitcoin Hashrate

The Hashrate of the Bitcoin network provides a useful insight into the growth of financially-motivated expenditure of computing resources to compute the solution of a single problem. I suspect that this will be a useful point of comparison as the network's exercised capacity approaches levels comparable to the the human brain.

```{r}
ai_bitcoin_hashrate %>%
  ggplot(aes(Date, GHPS)) +
  geom_line()
```

### Animal Brains

As our computational capacity climbs through the ranks of the animal kingdom, there are a variety of metrics which would be useful for comparison. The [Number of Neurons, Cortical Neurons (in mammals), Synapses](https://en.wikipedia.org/wiki/List_of_animals_by_number_of_neurons), [Brain size](https://en.wikipedia.org/wiki/Brain_size), [Brain-to-Body Mass Ratio](https://en.wikipedia.org/wiki/Brain-to-body_mass_ratio), [Encephalization Quotient](https://en.wikipedia.org/wiki/Encephalization_quotient) and [Cranial Capacity](https://en.wikipedia.org/wiki/Brain_size#Cranial_capacity) might all be useful in this line of research. Sadly, I've not yet found any sources (let alone reliable ones) for more than a few species. This dataset was scraped from the [Wikipedia's List of Animals by Number of Neurons](https://en.wikipedia.org/w/index.php?title=List_of_animals_by_number_of_neurons&oldid=710786267).

```{r}
ai_animal_neurons %>%
  ggplot(aes(Neurons, Synapses)) +
  geom_point() +
  scale_y_log10() +
  scale_x_log10() +
  stat_smooth(method = "lm")
```

The logarithmic scale in both dimensions suggests a [power-law relationship](https://en.wikipedia.org/wiki/Power_law), but this is derived from a very small, very noisy sample.

### Future of Life Institute Winning Grants

I don't know if there's anything interesting to be inferred from the [Future of Life Institute Grant Recipients](http://futureoflife.org/first-ai-grant-recipients/), but I collected this data when it was first published and this seems as appropriate a venue as any for its dissemination.

```{r}
ai_fli_winners %>%
  group_by(Institution) %>%
  summarise(Total = sum(Amount)) %>% 
  ggplot(aes(Institution, Total)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle=90, hjust = 1))
```

## Desired Data

This notably misses estimates about the computational capacities of the biological brains. Other types of estimates about the growth of computing power (e.g. global computing power, specs on the most powerful supercomputer at any given time, size of various commercial cloud infrastructure providers, etc.) are also needed. If you know of any reliable sources of these or related to these topics, please [email me](anthony@boyles.cc).

## References
